Plug-and-play dual-tree algorithm runtime analysis

### 2014-12-02 bounds.tex
The appendix is embarrassingly wrong.

\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
Numerous machine learning algorithms contain pairwise statistical problems in
their core. Often, tree structures are used to solve these problems efficiently.
Dual-tree algorithms using cover trees have rigorous worst-case runtime
guarantees for some of these problems. In this paper, we present a {\em
problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
the cover tree, separating out the problem-dependent and the problem-independent
elements. This allows us to just plug in bounds for the problem-dependent
elements to get runtime guarantees for dual-tree algorithms for any pairwise
statistical problem. We demonstrate this plug-and-play procedure for
nearest-neighbor search and approximate kernel density estimation to get
improved runtime guarantees.  Under some assumptions, we also present the first
linear runtime guarantee for dual-tree based range search.
\end{abstract}

@@

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.  We
then investigate the empirical performance of cover trees for these algorithms
in Section \ref{sec:emp}, with a focus on understanding the dataset-dependent
constants that our bounds use.

@@

\section{Empirical investigations}
\label{sec:emp}

\input{bounds-files/empirical.tex}

\section{Conclusion}

@@

% Appendix doesn't have anything useful or non-straightforward.
\appendix
\input{appendix-ec}

>>>>

\begin{abstract}
%A large collection of machine learning algorithms contain subproblems that can
%be efficiently solved with trees.  In the situation where a dual-tree algorithm
%with cover trees is used to solve a problem, rigorous worst-case runtime bounds
%can often be shown.  A worst-case runtime bound is proven for all cover tree
%dual-tree algorithms; this bound can be employed to give useful runtime bounds
%for individual problems.  We show a new, tighter bound of $O(c^9 \nu N)$ for
%all-nearest-neighbor search and also provide linear runtime bounds for
%approximate kernel density estimation and exact range count/search under some
%mild assumptions.  These are the first linear runtime bounds we are aware of for
%range count/search.
Numerous machine learning algorithms contain pairwise statistical problems at
their core. Often, tree structures are used to solve these problems efficiently.
Dual-tree algorithms using cover trees have rigorous worst-case runtime
guarantees for some of these problems. In this paper, we present a {\em
problem-independent} runtime guarantee for {\em any} dual-tree algorithm using
the cover tree, separating out the problem-dependent and the problem-independent
elements. This allows us to just plug in bounds for the problem-dependent
elements to get runtime guarantees for dual-tree algorithms for any pairwise
statistical problem. We demonstrate this plug-and-play procedure for
nearest-neighbor search and approximate kernel density estimation to get
improved runtime guarantees.  Under some assumptions, we also present the first
linear runtime guarantee for dual-tree based range search.
\end{abstract}

@@

Section \ref{sec:preliminaries} lays out the required background, notation, and
introduces the cover tree and its associated theoretical properties.  Readers
familiar with the cover tree literature and dual-tree algorithms
\citep[especially][]{curtin2013tree} may find that section to be review.
Following that, we introduce an intuitive measure of cover tree imbalance in
Section \ref{sec:imbalance}.  This measure of imbalance is then used to prove
the main result of the paper in Section \ref{sec:bound}, which is a worst-case
runtime bound for generalized dual-tree algorithms.  We apply this result to
three specific problems: nearest neighbor search (Section \ref{sec:nns}),
approximate kernel density estimation (Section \ref{sec:akde}), and range search
/ range count (Section \ref{sec:rs}), showing linear runtime bounds for each of
those algorithms.  Each of these bounds is an improvement on the
state-of-the-art, and in the case of range search, is the first such bound.
%  We
%then investigate the empirical performance of cover trees for these algorithms
%in Section \ref{sec:emp}, with a focus on understanding the dataset-dependent
%constants that our bounds use.

@@

%\section{Empirical investigations}
%\label{sec:emp}

%\input{bounds-files/empirical.tex}

\section{Conclusion}
\label{sec:conclusion}

@@

% Appendix doesn't have anything useful or non-straightforward.
%\appendix
%\input{appendix-ec}
